---
title: "Cross-sectional analysis of self-promotional language in texts of grant proposals across gender"
header-includes:
   - \usepackage{dcolumn}
   - \usepackage{booktabs}
   - \usepackage{makecell}
output:
  bookdown::word_document2:
    fig_caption: true
    keep_md: true
urlcolor: #6684c1
bibliography: bib_bmjopen.bib
link-citations: true
---


```{r init, include=FALSE, warning=FALSE}

###  PF, CF, Spark reorder the tables
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      message = FALSE)
knitr::opts_chunk$set(fig.showtext = TRUE)
knitr::knit_hooks$set(
  document = function(x) {
    sub("\\usepackage[]{color}",
        "\\usepackage[table]{xcolor}",
        x,
        fixed = TRUE)
  }
)
options(tinytex.clean = T)

# kableExtra for kable-tables
options(kableExtra.latex.load_packages = FALSE )
library(kableExtra)

library(snf.datastory)
# devtools::install_github("snsf-data/snf.datastory")
# or renv::install("snsf-data/snf.datastory")

# Use here("...") with paths relative to the project root
# to address files
library(here)
# Avoid dependency on load order of packages
# library(conflicted)

# conflict_prefers 
# if you'd like to see futher updates, tell rhe.
# conflict_prefer("filter", "dplyr")
# conflict_prefer("select", "dplyr")
# conflict_prefer("here", "here")
# conflict_prefer("list", "base")
# conflict_prefer("as_image", "kableExtra")
# conflict_prefer("group_rows", "kableExtra")
# More libraries
library(cowplot)
library(texreg)
library(broom)
library(dplyr)
library(tm)
# library(cld3)
library(sandwich)
library(msm)
library(tidyr)
library(ggpubr)
library(flextable)
library(ftExtra)
library(stringr)


library(readxl)
# windowsFonts("Source Sans" = windowsFont("Source Sans Pro Regular"))
# change this to false if desired output is pdf
is_word = TRUE
```

```{r load-ls-data, cache = FALSE, echo = FALSE, results = 'hide'}

# Load the data from Zenodo:
url <- "https://zenodo.org/record/6676562/files/positive_words_frequencies.xlsx"
destfile <- "positive_words_frequencies.xlsx"
curl::curl_download(url, destfile)

career_freq_ls <-  read_xlsx(destfile, sheet = "career") %>% 
  mutate(Number = 1:n()) %>% 
  mutate(ResearchInstitutionType = 
           case_when(ResearchInstitutionType %in% 
                       c("Other", "Other research institutes", 
                         "Hospitals (non-university)",
                         "University of applied sciences") ~ "UAS and Other",
                     TRUE ~ ResearchInstitutionType))

sparkabs_freq_ls <- read_xlsx(destfile, sheet = "spark") %>% 
  mutate(Number = 1:n()) %>% 
  mutate(ResearchInstitutionType = 
           case_when(ResearchInstitutionType %in% 
                       c("Other", "Other research institutes", 
                         "Hospitals (non-university)",
                         "University of applied sciences") ~ "UAS and Other",
                     TRUE ~ ResearchInstitutionType))

pf_freq_ls <- read_xlsx(destfile, sheet = "projects") %>% 
  mutate(Number = 1:n()) %>% 
  mutate(ResearchInstitutionType = 
           case_when(ResearchInstitutionType %in% 
                       c("Other", "Other research institutes", 
                         "Hospitals (non-university)",
                         "University of applied sciences") ~ "UAS and Other",
                     TRUE ~ ResearchInstitutionType))

# 25 positive words as suggested in Vinkers et al (2015)
pos_words <- c(
  "novel",
  "unique",
  "promising",
  "favorable|favourable",
  "robust",
  "excellent",
  "prominent",
  "supportive",
  "encouraging",
  "remarkable",
  "innovative",
  "unprecedented",
  "bright",
  "enormous",
  "reassuring",
  "creative",
  "assuring",
  "hopeful",
  "astonishing",
  "spectacular",
  "amazing",
  "inventive",
  "phenomenal",
  "groundbreaking",
  "inspiring"
)

```


```{r load-function-etc, cache = FALSE, echo = FALSE, results = 'hide'}
# Code to format tables and figures
source(here::here("r_functions.R"))
```


```{r load-function-etc2, cache = FALSE, echo = FALSE, results = 'hide'}
# Number of proposals
total_spark <- nrow(sparkabs_freq_ls)
total_pf <- nrow(pf_freq_ls)
total_career <- nrow(career_freq_ls)

# Percentage international applicants
pct_int_spark <- table(sparkabs_freq_ls$NationalityIsoCode=="CH")[1]/
  table(is.na(sparkabs_freq_ls$NationalityIsoCode))[1]*100
pct_int_pf <- table(pf_freq_ls$NationalityIsoCode=="CH")[1]/
  table(is.na(pf_freq_ls$NationalityIsoCode))[1]*100
pct_int_career <- table(career_freq_ls$NationalityIsoCode=="CH")[1]/
  table(is.na(career_freq_ls$NationalityIsoCode))[1]*100

# Success Rates
sr_spark <- sparkabs_freq_ls %>%
  summarise(sr = round(length(which(IsFundable == 1)) / n() * 100, 1))
sr_pf <- pf_freq_ls %>%
  summarise(sr = round(length(which(IsApproved == 1)) / n() * 100, 1))
sr_career <- career_freq_ls %>% 
  summarise(sr = round(length(which(IsApproved == 1)) / n() * 100, 1))


## ----------------------------
## data summary
## ----------------------------
# Gender distribution of applicants and success rate per research area
# ... spark abstracts (only look at fundable vs non-fundable)
gender_spark <- sparkabs_freq_ls %>%
  group_by(ResponsibleApplicantGender,
           IsFundable) %>%
  count() %>%
  group_by(ResponsibleApplicantGender) %>%
  mutate(total = sum(n),
         rate_plot = n/total,
         rate = paste0(format(round(n * 100 / total, 1), nsmall = 1), "%")) %>%
  # filter(IsFundable == 1) %>%
  select(ResponsibleApplicantGender,
         IsFundable,
         n,
         rate,
         rate_plot,
         total)


# ... pf
gender_pf <- pf_freq_ls %>%
  group_by(ResponsibleApplicantGender,
           IsApproved) %>%
  count() %>% 
  group_by(ResponsibleApplicantGender) %>%
  mutate(total = sum(n),
         rate_plot = n/total,
         rate = paste0(format(round(n * 100 / total, 1), nsmall = 1), "%")) %>%
  # filter(IsApproved == 1) %>%
  select(ResponsibleApplicantGender,
         IsApproved,
         n,
         rate,
         rate_plot,
         total)

# ... career
gender_career <- career_freq_ls %>%
  group_by(ResponsibleApplicantGender,
           IsApproved) %>%
  count() %>%
  group_by(ResponsibleApplicantGender) %>%
  mutate(total = sum(n),
         rate_plot = n/total,
         rate = paste0(format(round(n * 100 / total, 1), nsmall = 1), "%")) %>%
  # filter(IsApproved == 1) %>%
  select(ResponsibleApplicantGender,
         IsApproved,
         n, 
         rate,
         rate_plot,
         total)


# EDA % of applications with positive words

spark_numbers <- sparkabs_freq_ls %>% 
  select(Number, sum_pos) %>%
  mutate(is_positive = if_else(sum_pos > 0, 1, 0)) %>%
  group_by(is_positive) %>%
  summarise(n_app = n(), .groups = "drop")

pf_numbers <- pf_freq_ls %>% 
  select(Number, sum_pos) %>%
  mutate(is_positive = if_else(sum_pos > 0,1,0)) %>%
  group_by(is_positive) %>%
  summarise(n_app = n(), .groups = "drop")

career_numbers <- career_freq_ls %>% 
  select(Number, sum_pos) %>%
  mutate(is_positive = if_else(sum_pos > 0, 1, 0)) %>%
  group_by(is_positive) %>%
  summarise(n_app = n(), .groups = "drop")


# EDA overall success rate
spark_success <- sparkabs_freq_ls %>% 
  group_by(IsFundable) %>%
  summarise(n_app= n(), .groups = "drop")

pf_success <- pf_freq_ls %>% 
  group_by(IsApproved) %>%
  summarise(n_app = n(), .groups = "drop")

career_success <- career_freq_ls %>% 
  group_by(IsApproved) %>%
  summarise(n_app = n(), .groups = "drop")

## -----------------------------------------------
## analyze specific words
## -----------------------------------------------

## ----plot specific words ----
sparkabs_plot <-
  plot_positivewords(freq_word = sparkabs_freq_ls, 
                     pos_words= pos_words)
pf_plot <- plot_positivewords(freq_word = pf_freq_ls,
                              pos_words = pos_words)
career_plot <-
  plot_positivewords(freq_word = career_freq_ls,
                     pos_words = pos_words)
```

```{r all calculations part 2, cache = FALSE, echo = FALSE, results = 'hide'}
#analyze the effects of gender on language
#--------------------------------------
## analysis: gender on language
#--------------------------------------
# Analysis of the whether gender has an effect on the use of words as through..
# .. binary variable for the presence of at least one positive word
# .. .. using linear probability models
# .. .. using logistic regression models
# .. the count of distinct positive words using Poisson regression
# .. .. quasipoisson and negative binomial regression


# spark abstracts
# ---------------
sparkabs_reg_ls <- sparkabs_freq_ls %>%
  analyze_gender()
# pf abstracts
# ---------------
pf_reg_ls <- analyze_gender(dat = pf_freq_ls)

# career abstracts
# ---------------
career_reg_ls <- analyze_gender(dat =  career_freq_ls)

```



# Abstract{-}
**Objectives:**To investigate whether the use of positive, self-promotional language in grant proposals correlates with the gender of the applicant. Different types of funding schemes are examined.

**Setting:**  8,150 grant proposals submitted to the Swiss National Science Foundation (SNSF) in life sciences disciplines across three different funding schemes, targeting grant applicants at different career stages.

**Design:** Cross-sectional analysis using multivariate, logistic, and Poisson regression models.

**Participants:** Grant proposals, including demographic data about their applicants. 

**Primary outcome measure:** Count and binary presence of positive words, as identified in previous work analysing positive language usage in published biomedical articles, in grant proposal abstracts. 

**Results:** No strong evidence of an association between the use of self-promotional language and gender were identified in grant proposals from life sciences disciplines. In the Projects funding scheme that funds independent bottom-up research, male applicants have used on average slightly more positive words in their proposal abstracts than their female peers, but the estimates (Rate Ratio of 1.044; 95% CI 0.987 to 1.104) show no evidence of an effect. Similarly, no evidence of gender differences was found in proposals from the Careers funding schemes (Rate Ratio of 1.077, 95% CI 0.992–1.17) which target early career researchers, and in Spark (Rate Ratio of 1.093, 95% CI 0.953–1.254), a double-blind funding scheme supporting unconventional research. 

**Conclusion(s):** The goal of this study is to examine whether there is a difference in positive language usage between male and female grant applicants. It is encouraging that there is no strong evidence indicating a difference in the use of positive words between male and female grant applicants. Although many factors influence how peer reviewers perceive the content of research grant proposals, language usage undoubtedly plays an important role. We urge other funding agencies to conduct similar studies to identify if language usage contributes to biases in grant evaluation. 


**Strengths and Limitations**

* This study is an extension of previous work examining published biomedical articles, and is here applied to the new setting of grant proposals. To our knowledge, this is the first study of this type examining research grant proposals, and includes a large sample of both funded and rejected research grant proposals. 

* This study investigates the counts of positive words used in grant proposals through Poisson regression models, a method not previously applied in this setting.

* This study does not assess whether the use of self-promotional language contributes to grant funding success. 

* This study investigated the presence of 25 terms previously found to be associated with self-promotional language. Many other dimensions of language usage may influence how readers perceive the content of research grant proposals, but this is beyond the scope of the current work. 


# Introduction

Bias against female grant applicants has been observed across research funding agencies. A study from the Swiss National Science Foundation found that both male and female peer reviewers graded female applicants more harshly than male applicants, even when potentially confounding factors were taken into consideration (@Severin2020). In the US, a text analysis by the National Institutes of Health found that peer reviewers used different language to describe male and female grant applicants, where male applicants were more likely to be described as leaders (@Magua2017). The Canadian Institutes of Health Research found that male applicants were more likely to receive funding than female applicants, but that this was because female applicants were evaluated less favorably as principal investigators, not because of any difference in the quality of their proposals (@Witteman2019). Evidence of bias in the evaluation of track records of female applicants was also found in the Netherlands (@Van_der_Lee2015). The current study aims to determine if the language used by the applicants themselves in their abstracts might be a contributing factor in the differing treatment of grant proposals from male and female principal investigators. 

It is well-established that scientists often over-inflate the importance of their work in order to get published or funded (@Ioannidis2014, @Chiu2017). There is evidence that men use more self-promotional language than women when describing their own work histories (@Reuben2014). In particular, female authors of medical and life sciences articles were previously observed to use strongly positive language less frequently to describe their research than their male counterparts (@Lerchenmueller).  @Lerchenmueller used a list of 25 positive words established by @Vinkers2015 who selected them after manual analysis of biomedical texts from PubMed. @Vinkers2015 observed a significant increase in the use of these words across all PubMed literature over a 40-year period, and concluded that this is likely due to the positive-outcome bias in scientific publishing - that only research with the most unprecedented positive results will be published. This list was developed specifically for studying biomedical texts, and preliminary analyses have suggested that the usage of these words differs somewhat in other research areas.  

The current study builds on this study by @Lerchenmueller, replicating and extending their analyses to understand if their findings can also be observed in the context of research grant proposals. We analyse proposals submitted to three different funding schemes at the Swiss National Science Foundation, where the importance of the track record of the applicant varies considerably. We hypothesize that whether an applicant must present themselves as an established and successful researcher, as opposed to just promoting a single project idea, may influence how they describe their research. Since writing styles can differ substantially between academic disciplines (@Hyland2006) and the list of positive words was developed for biomedical texts, we will limit our analysis to grant proposals coming from the broad domain of life sciences. 
This study provides insight into the use of self-promotional language in research grant proposals. The results should be of broad interest to the international scientific community, as Swiss scientific research is very competitive internationally. In fact, over half of the Swiss-based researchers included in this study originated from countries outside Switzerland (based on their current citizenship). This analysis is also important as a replication attempt of a notable result in the literature. 

# Methods

## Data: grant proposals

```{r summary-statistics}

tab_fi <- pf_freq_ls %>% 
  summarise(n_perc = format(n(), big.mark = "'")) %>% 
  bind_cols(career_freq_ls %>%
              summarise(n_perc_car = format(n(), big.mark = "'"))) %>% 
  bind_cols(sparkabs_freq_ls %>%
              summarise(n_perc_spark = format(n(), big.mark = "'"))) %>% 
  mutate(variable = "") %>% 
  select(variable, everything())

tab_gender <- pf_freq_ls %>% 
  group_by(ResponsibleApplicantGender) %>% 
  count() %>% 
  ungroup() %>% 
  mutate(perc = n / sum(n) * 100,
         n_perc = paste0(format(n, big.mark = "'"), " (", round(perc, 1), "%)"),
         ResponsibleApplicantGender = 
           case_when(ResponsibleApplicantGender == "f" ~ "Female applicants",
                     ResponsibleApplicantGender == "m" ~ "Male applicants")) %>% 
  rename(variable = ResponsibleApplicantGender) %>% 
  select(variable, n_perc) %>% 
  bind_cols(career_freq_ls %>% 
              group_by(ResponsibleApplicantGender) %>% 
              count() %>% 
              ungroup() %>% 
              mutate(perc = n / sum(n) * 100,
                     n_perc_car = paste0(format(n, big.mark = "'"), 
                                         " (", round(perc, 1), "%)")) %>% 
              select(n_perc_car)) %>% 
    bind_cols(sparkabs_freq_ls %>% 
              group_by(ResponsibleApplicantGender) %>% 
              count() %>% 
              ungroup() %>% 
              mutate(perc = n / sum(n) * 100,
                     n_perc_spark = paste0(format(n, big.mark = "'"),
                                           " (", round(perc, 1), "%)")) %>% 
                select(n_perc_spark))

tab_inst <- pf_freq_ls %>% 
  group_by(ResearchInstitutionType) %>% 
  count() %>% 
  ungroup() %>% 
  mutate(perc = n / sum(n) * 100,
         n_perc = paste0(format(n, big.mark = "'"),
                         " (", round(perc, 1), "%)")) %>% 
  rename(variable = ResearchInstitutionType) %>% 
  select(variable, n_perc) %>% 
  bind_cols(career_freq_ls %>%  
              group_by(ResearchInstitutionType) %>% 
              count() %>% 
              ungroup() %>% 
              mutate(perc = n / sum(n) * 100,
                     n_perc_car = paste0(format(n, big.mark = "'"), 
                                         " (", round(perc, 1), "%)")) %>% 
              select(n_perc_car)) %>% 
  bind_cols(sparkabs_freq_ls %>% 
              group_by(ResearchInstitutionType) %>% 
              count() %>% 
              ungroup() %>% 
              mutate(perc = n / sum(n) * 100,
                     n_perc_spark = paste0(format(n, big.mark = "'"),
                                           " (", round(perc, 1), "%)")) %>% 
                select(n_perc_spark))

tab_year <- pf_freq_ls %>%  
  group_by(CallYear) %>% 
  count() %>% 
  ungroup() %>% 
  mutate(perc = n / sum(n) * 100,
         n_perc = paste0(format(n, big.mark = "'"),
                         " (", round(perc, 1), "%)")) %>% 
  mutate(variable = as.character(CallYear)) %>% 
  select(variable, n_perc) %>% 
  left_join(career_freq_ls %>%   
              group_by(CallYear) %>% 
              count() %>% 
              ungroup() %>% 
              mutate(perc = n / sum(n) * 100,
                     n_perc_car = paste0(format(n, big.mark = "'"), 
                                         " (", round(perc, 1), "%)")) %>% 
              mutate(variable = as.character(CallYear)) %>%
              select(variable, n_perc_car)) %>% 
  left_join(sparkabs_freq_ls %>% 
              group_by(CallDecisionYear) %>% 
              count() %>% 
              ungroup() %>% 
              mutate(perc = n / sum(n) * 100,
                     n_perc_spark = paste0(format(n, big.mark = "'"),
                                           " (", round(perc, 1), "%)")) %>% 
              mutate(variable = as.character(CallDecisionYear)) %>% 
              select(variable, n_perc_spark))

tab_age <- pf_freq_ls %>%  
  summarise(n_perc = 
              paste0(round(mean(ResponsibleApplicantAgeAtSubmission), 1), 
                     " (", min(ResponsibleApplicantAgeAtSubmission), " - ",
                     max(ResponsibleApplicantAgeAtSubmission), ")")) %>% 
  mutate(variable = "Mean (min-max)") %>% 
  select(variable, n_perc) %>% 
  bind_cols(career_freq_ls %>% 
              summarise(
                n_perc_car = 
                  paste0(round(mean(ResponsibleApplicantAgeAtSubmission), 1), 
                         " (", min(ResponsibleApplicantAgeAtSubmission), " - ",
                         max(ResponsibleApplicantAgeAtSubmission), ")")) %>% 
              select(n_perc_car)) %>% 
  bind_cols(sparkabs_freq_ls %>% 
              summarise(
                n_perc_spark = 
                  paste0(round(mean(ResponsibleApplicantAgeAtSubmission), 1), 
                         " (", min(ResponsibleApplicantAgeAtSubmission), " - ",
                         max(ResponsibleApplicantAgeAtSubmission), ")")) %>% 
              select(n_perc_spark))

tab_textlength <- pf_freq_ls %>%  
  summarise(n_perc = paste0(round(mean(text_length), 1), " (",
                            min(text_length), " - ",
                            format(max(text_length), big.mark = "'"), ")")) %>% 
  mutate(variable = "Mean (min - max)") %>% 
  select(variable, n_perc) %>% 
  bind_cols(career_freq_ls %>% 
              summarise(
                n_perc_car =
                  paste0(round(mean(text_length), 1), " (",
                         min(text_length), " - ", 
                         format(max(text_length), big.mark = "'"), ")")) %>% 
              select(n_perc_car)) %>% 
  bind_cols(sparkabs_freq_ls %>% 
              summarise(
                n_perc_spark = 
                  paste0(round(mean(text_length), 1), " (",
                         min(text_length), " - ", 
                         format(max(text_length), big.mark = "'"), ")")) %>% 
              select(n_perc_spark))

tab_sum_pos <- pf_freq_ls %>%  
  summarise(n_perc = paste0(round(mean(sum_pos), 1), " (", min(sum_pos), " - ",
                            max(sum_pos), ")")) %>% 
  mutate(variable = "Mean (min - max)") %>% 
  select(variable, n_perc) %>% 
  bind_cols(career_freq_ls %>% 
              summarise(
                n_perc_car = paste0(round(mean(sum_pos), 1), " (", 
                                    min(sum_pos), "-", max(sum_pos), ")")) %>% 
              select(n_perc_car)) %>% 
  bind_cols(sparkabs_freq_ls %>% 
              summarise(
                n_perc_spark = paste0(round(mean(sum_pos), 1), " (",
                                      min(sum_pos), "-", max(sum_pos), ")")) %>% 
              select(n_perc_spark))
```


The data comes from research grant proposals submitted to the Swiss National Science Foundation (SNSF) in life sciences disciplines in the ten years between 2011 and 2020, and one shorter funding scheme. This study looks at the titles and abstracts from three different funding schemes and investigates whether any of the 25 positive words identified by @Vinkers2015 appear in these texts, and if so, how frequently. Only English-language proposals are considered, which means that 0.3% of the available Career proposals and 0.3% of the Project proposals had to be removed before processing.  

As part of the proposal submission process, applicants identify their own gender, so that the gender of each applicant is known with certainty. For the purposes of this study, we only consider the gender of the corresponding applicant on each proposal, although one of the three funding schemes included (Projects) can also involve multiple co-applicants.

We investigate whether there is an association between the gender and the use of positive words in each funding scheme. Grant proposals to the SNSF are classified into three research areas: 1) Social Sciences and Humanities, 2) Mathematics, Natural- and Engineering Sciences, and 3) Life Sciences. Because the list of positive terms considered was created for biomedical texts, and we do not knowit is uncertain how well this list generalizes to other domains, we include only life sciences proposals in this analysis. 

Researchers apply from three types of research institutions: Cantonal universities are the 12 large public research universities in Switzerland; the ETH-Domain consists of the Swiss Federal Institutes of Technology (two universities - ETHZ and EPFL - and four research institutions); and the Universities of Applied Sciences (UAS) and other research institutes focused on practice-oriented qualifications.

The three funding schemes included in this study are:

**Projects**: This is the biggest funding scheme at the SNSF, and funds bottom-up research. Independent researchers with contracts at Swiss research institutions are eligible to apply for this funding scheme. The track record is considered, but is only one part of the complete proposal. The researchers are on average `r tab_age %>% select(n_perc) %>% substr( 1, 4)` years old and most of them are employed at cantonal universities, `r tab_inst %>% filter(variable == "Cantonal university") %>% pull(n_perc)`. 
<!-- 15,785  -->
`r pf_freq_ls %>% nrow() %>% format(big.mark = "'")` proposals from Projects are included in this study, 
<!-- 51.7 -->
`r pf_numbers %>% mutate(perc = n_app/sum(n_app)) %>% filter(is_positive == 1) %>% pull(perc) %>% round(2) *100`% of which contain at least one of the positive words in their title or abstract. 

**Careers** (Ambizione, Excellenza, Professorships): These funding schemes are meant to support early-career researchers to be able to lead independent projects. The target researchers are a couple years past the completion of their PhDs, and are aiming for, but do not yet have, permanent tenure-track professorships. The track record of the applicants is very important for the success of their proposals. 
<!-- 4,648  --> 
`r career_freq_ls %>% nrow() %>% format(big.mark = "'")`
proposals from Careers are included in this study, 
<!-- 55.1 -->
`r career_numbers %>% mutate(perc = n_app/sum(n_app)) %>% filter(is_positive == 1) %>% pull(perc) %>% round(2) *100`% of which contain at least one of the positive words in their title or abstract. Compared to the Projects instrument, the applicants to the career schemes are on average younger (`r tab_age %>% select(n_perc_car) %>% substr( 1, 4)` years). 

**Spark**: This was launched as a pilot funding scheme that is meant to support rapid funding of unconventional ideas. The pilot consisted of two calls in 2020 and this dataset is smaller than the other two. The evaluation procedure is double-blind, and so the track record of the applicant should be of no importance to the proposal. 
<!-- 1,532  -->
`r sparkabs_freq_ls%>% nrow() %>% format(big.mark = "'")` proposals from Spark are included in this study, 
<!-- 61.2  -->
`r spark_numbers %>% mutate(perc = n_app/sum(n_app)) %>% filter(is_positive == 1) %>% pull(perc) %>% round(2) *100`% of which contain at least one of the positive words in their title or abstract. Compared to the other two funding schemes, the applicants to Spark are employed at the Universities of Applied Sciences at a higher percentage `r  (tab_inst %>% filter(variable == "UAS and Other") %>% pull(n_perc_spark) %>% str_extract_all("\\([^()]*\\)|[^()]+"))[[1]][2]`.
<!-- We note that the available funding was very different between the two Spark calls, so when we analyse successful proposals, we will consider "fundability" for Spark, rather than true funding success, in order to be able to use the same success criterion for both calls.  -->
The applicants were on average `r tab_age %>% select(n_perc_spark) %>% substr( 1, 4)` years old.


## Data: Positive Words

The positive words included in this study, as defined by @Vinkers2015 based on publications from PubMed, are: amazing, assuring, astonishing, bright, creative, encouraging, enormous, excellent, favorable/ favourable, groundbreaking, hopeful, innovative, inspiring, inventive, novel, phenomenal, prominent, promising, reassuring, remarkable, robust, spectacular, supportive, unique, and unprecedented.

## Statistical analysis

To determine whether the use of positive language is associated with the gender of the grant applicant, @Lerchenmueller used multivariate linear probability regressions to model the binary outcome of whether a positive word was used at least once. To replicate their results, we followed the same procedure. In addition, we perform a logistic regression to model the same binary outcome, as we view this type of model to be more suited to answer our research question.  In addition, we perform a Poisson regression to model the count of positive word, to see if one gender might use a higher number of positive words. For all models, the age of the applicant, the type of institution, and the length of the text are taken as covariates, with gender as the variable of interest. 
The models are adjusted for age because we consider the applicant’s age to be a confounder. Age is a reasonable proxy for career stage, and it is understandable that researchers at different stages of their careers would construct texts differently. We additionally adjust for the institution type as a reasonable, but imperfect proxy for prestige. Depending on the reputation of the institutions, as well as the investment the institutions make in their individual grant writing services, the affiliated researchers may write more positively about their research (for example, see Figure S1 in the [online supplement](https://snsf-data.github.io/positive_language/online_supplement/index.html)). The gender distributions at the different institution types varies considerably. The models are adjusted for the length of the texts (in word count per 100 words), as longer texts have higher likelihood of containing more positive words. Proposals to Projects, Careers and Spark are analysed in separate models. 

We will determine the presence or absence of evidence of an effect based on effect estimates with 95% confidence intervals, which address the size of an effect more directly than p-values (@Wasserstein2016). Odds ratio and rate ratio estimates will be used for the logistic and the Poisson regression models respectively. All analyses were done in R [REF] version 4.3.2. 

<!-- \pagebreak -->


# Results
## Summary Statistics

Data on the number and distribution of proposals submitted to the three investigated funding schemes in the Life Sciences is provided in Table \@ref(tab:tab1). 

```{r tab1, message = FALSE, warning = FALSE, cache = F, results = "asis"}

tab1 <- tab_fi %>% 
  mutate(group = "Total") %>% 
  bind_rows(tab_gender %>% 
              mutate(group = "Gender"),
            tab_inst %>% 
              mutate(group = "Institution Type"), 
            tab_age %>% 
              mutate(group = "Age at submission"),
            tab_textlength %>% 
              mutate(group = "Text length"), 
            tab_sum_pos %>% 
              mutate(group = "Number positive words"),
            tab_year %>% 
              mutate(group = "Call Year")) %>% 
  rename(" " = variable, 
         "Projects" = n_perc, 
         "Careers" = n_perc_car,
         "Spark" = n_perc_spark) %>% 
  as_grouped_data(groups = c("group")) %>% 
  as_flextable(hide_grouplabel = TRUE) %>% 
  set_caption(caption = "Descriptive Statistics for the three funding schemes (Projects, Careers and Spark): count and percentage for factor variables, mean, maximum and minimum for continuous variables.") %>% 
  hline(i = c(2, 5, 9, 11, 13, 15)) %>% 
  bold(i = c(1, 3, 6, 10, 12, 14, 16)) %>% 
  bold(part = "header") %>% 
  width(width = 1.8) %>% 
  fontsize(size = 9, part = "body")

tab1 
```


The percentage of grant proposal from the funding scheme Projects that include any of the positive words in their title or abstract per Call year and gender of the applicant is shown in Figure \@ref(fig:perc-one-poswords). We observe that the use of positive terms by male applicants has been increasing over time.


```{r perc-one-poswords, fig.height = 5, fig.align = "center", fig.width=7, fig.cap = "The share of proposals from Projects that include at least one of the defined positive words in their title or abstract.", message = FALSE, warning = FALSE}
pf_freq_ls %>% 
  mutate(ResponsibleApplicantGender = 
           case_when(ResponsibleApplicantGender == "f" ~ "Female applicants",
                     ResponsibleApplicantGender == "m" ~ "Male applicants")) %>% 
  group_by(ResponsibleApplicantGender, CallYear) %>% 
  count(test = sum_pos > 0) %>% 
  mutate(freq = n/sum(n)) %>% 
  filter(test) %>% 
  ggplot(aes(x = as.factor(CallYear), y = freq, 
             fill = ResponsibleApplicantGender)) +
  geom_col(position = position_dodge(), 
           color = "white", alpha = .8) + 
  scale_fill_manual(
    values = get_datastory_scheme(),
    name = "",
    labels = c("Female", "Male")) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1), 
                     limits=c(0, .7)) +
  labs(y = "Share of at least one positive word", x = "Call year") +
  theme_minimal() + 
  theme(panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line(linetype = "dashed"),
        panel.grid.minor.y = element_line(linetype = "dashed")) 
```

The distributions of the positive words across all three funding schemes are provided in Figure S2 included in the [online supplement](https://snsf-data.github.io/positive_language/online_supplement/index.html). Figure \@ref(fig:fig-distribution-number-poswords) shows the distribution of the number of positive words per funding scheme and gender of the corresponding applicant. We see that the majority of proposals include at least one positive term in their title or abstract.

```{r fig-distribution-number-poswords, fig.align="center", fig.height = 8, fig.width = 6, fig.cap = "The distribution of the number of positive words used in the abstracts of the different funding schemes by gender of the corresponding applicant. All counts larger than 7 are regrouped into 7+.", message = FALSE, warning = FALSE}

p_pf <- pf_freq_ls %>% 
  group_by(ResponsibleApplicantGender) %>% 
  count(sum_pos) %>% 
  mutate(sum_pos = case_when(sum_pos >= 7 ~ "7+",
                             TRUE ~ as.character(sum_pos)),
         sum_pos = factor(sum_pos, levels = c(0:6, "7+"))) %>% 
  group_by(ResponsibleApplicantGender, sum_pos) %>% 
  summarise(n = sum(n)) %>%
  mutate(freq = n/sum(n)) %>% 
  ggplot(aes(x = sum_pos, y = freq, color = ResponsibleApplicantGender,
             fill = ResponsibleApplicantGender)) +
  geom_col(position = position_dodge(), color = "white", alpha = 0.8) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1), 
                     limits=c(0, .5)) +
  scale_fill_manual(
    values = get_datastory_scheme(),
    name = "",
    labels = c("Female", "Male")) +
  theme_minimal() + 
  labs(x = "Number of positive words", 
       y = "Share") +
  theme(legend.position = "bottom", 
        axis.title.x = element_blank(), 
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line(linetype = "dashed"),
        panel.grid.minor.y = element_line(linetype = "dashed")) +
  ggtitle(paste0("Projects")) 

p_careers <- career_freq_ls %>% 
  group_by(ResponsibleApplicantGender) %>% 
  count(sum_pos) %>% 
  mutate(sum_pos = case_when(sum_pos >= 7 ~ "7+",
                             TRUE ~ as.character(sum_pos)),
         sum_pos = factor(sum_pos, levels = c(0:6, "7+")))  %>% 
  group_by(ResponsibleApplicantGender, sum_pos) %>% 
  summarise(n = sum(n)) %>% 
  mutate(freq = n/sum(n)) %>% 
  ggplot(aes(x = sum_pos, y = freq, color = ResponsibleApplicantGender,
             fill = ResponsibleApplicantGender)) +
  geom_col(position = position_dodge(), color = "white", alpha = 0.8) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1), 
                     limits=c(0, .5)) +
  scale_fill_manual(
    values = get_datastory_scheme(),
    name = "",
    labels = c("Female", "Male")) +
  theme_minimal() + 
  labs(x = "Number of positive words", 
       y = "Share") +
  theme(axis.title.x = element_blank(), 
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line(linetype = "dashed"),
        panel.grid.minor.y = element_line(linetype = "dashed")) +
  ggtitle(paste0("Careers")) 

p_spark <- sparkabs_freq_ls %>% 
  group_by(ResponsibleApplicantGender) %>% 
  count(sum_pos) %>% 
  mutate(sum_pos = case_when(sum_pos >= 7 ~ "7+",
                             TRUE ~ as.character(sum_pos)),
         sum_pos = factor(sum_pos, levels = c(0:6, "7+"))) %>% 
  group_by(ResponsibleApplicantGender, sum_pos) %>% 
  summarise(n = sum(n)) %>% 
  mutate(freq = n/sum(n)) %>% 
  ggplot(aes(x = sum_pos, y = freq, color = ResponsibleApplicantGender,
             fill = ResponsibleApplicantGender)) +
  geom_col(position = position_dodge(), color = "white", alpha = 0.8) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1), 
                     limits=c(0, .5)) +
  scale_fill_manual(
    values = get_datastory_scheme(),
    name = "",
    labels = c("Female", "Male")) +
  theme_minimal() + 
  labs(x = "Number of positive words", 
       y = "Share") +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line(linetype = "dashed"),
        panel.grid.minor.y = element_line(linetype = "dashed")) +
  ggtitle(paste0("Spark")) 

p <- cowplot::plot_grid(p_pf + theme(legend.position = "none"), 
                        p_careers + theme(legend.position = "none"),
                        p_spark + theme(legend.position = "none"),
                        ncol = 1)
                        

grobs <- ggplotGrob(p_pf)$grobs
legend <-
  grobs[[which(sapply(grobs, function(x)
    x$name) == "guide-box")]]

plot_grid(p,
          legend,
          ncol = 1,
          rel_heights = c(1, 0.05))
```


## Statistical analysis of grant proposals 

We present results of the regression models relating the gender of the applicant to the use of positive language. 

The logistic regression model modeling the binary outcome “at least one positive word” is summarised in the first column of Table \@ref(tab:tab-gender-binary-count-ls). A summary table for the multivariate linear probability model modeling the binary outcome in the same way as @Lerchenmueller, can be found in Tables S1 in the [online supplement](https://snsf-data.github.io/positive_language/online_supplement/index.html). 
In these models, the only estimates suggesting evidence for an effect was found to be the length of the texts, i.e., positive coefficient with 95% CI not including zero. In Projects only, researchers affiliated to the ETH Domain were found to have a higher probability of using at least one positive word when describing their research projects. No evidence for a gender effect on the presence of at least one positive word is found in any
of the binary outcome models.

```{r tab-gender-linear-stratify, message = FALSE, warning = FALSE, cache = F, results = "asis"}
# list of covariates to be reported
covariate_list <- names(coefficients(sparkabs_reg_ls$linear))[-1]
# regression results 
reg_gender_linear_ls <- list(pf = pf_reg_ls$linear, 
                             career = career_reg_ls$linear ,
                             spark = sparkabs_reg_ls$linear)

tabulate_lm_ci <- function(m, digits = 3){
  e <- summary(m)$coef[, "Estimate"]
  sd <- summary(m)$coef[, "Std. Error"]
  paste0(round(e, digits), " (", round(e - 1.96*sd, digits),
         "--", round(e + 1.96*sd, digits), ")") %>% 
    return()
}

ls_column_linear <- lapply(reg_gender_linear_ls,
                           FUN = tabulate_lm_ci) %>% unlist()

var_column <- rep(c("Intercept", "gender - male", "age (years)",
                    "inst. type - ETH Domain", 
                    "inst. type - UAS and Other",
                    "text length (per 100 words)"), 3)

FI_column <- c(
  rep(paste0("Projects: ", format(nrow(pf_reg_ls$linear$model), big.mark = "'"),
             " observations"), 6),
  rep(paste0("Careers: ", format(nrow(career_reg_ls$linear$model),
                                 big.mark = "'"), " observations"), 6), 
  rep(paste0("Spark: ", format(nrow(sparkabs_reg_ls$linear$model),
                               big.mark = "'"), " observations"), 6))

tab_linear <- data.frame(FI_column,
                         var_column, ls_column_linear,
                         stringsAsFactors = FALSE)
row.names(tab_linear) <- NULL

```


```{r tab-gender-binary-stratify, message = FALSE, warning = FALSE, cache = F, results = "asis"}
# list of covariates to be reported
covariate_list <- names(coefficients(sparkabs_reg_ls$log))[-1]
# regression results 
reg_gender_binary_ls <- list(pf = pf_reg_ls$log, 
                             career = career_reg_ls$log ,
                             spark = sparkabs_reg_ls$log)

tabulate_ci <- function(m, digits = 3){
  v <- cbind(exp(m$coefficients), exp(confint(m))) 
  paste0(round(v[ ,1], digits), " (", round(v[ , 2], digits), "--",
         round(v[ ,3], digits), ")") %>% 
    return()
}

ls_column <- lapply(reg_gender_binary_ls,
                    FUN = tabulate_ci) %>% unlist()

var_column <- rep(c("Intercept", "gender - male", "age (years)",
                    "inst. type - ETH Domain", 
                    "inst. type - UAS and Other",
                    "text length (per 100 words)"), 3)

FI_column <- c(
  rep(paste0("Projects: ", format(nrow(pf_reg_ls$log$model), big.mark = "'"),
             " observations"), 6),
  rep(paste0("Careers: ", format(nrow(career_reg_ls$log$model), big.mark = "'"),
             " observations"), 6), 
  rep(paste0("Spark: ", format(nrow(sparkabs_reg_ls$log$model), big.mark = "'"),
             " observations"), 6))

tab_binary <- data.frame(FI_column,
                         var_column, ls_column,
                         stringsAsFactors = FALSE)
row.names(tab_binary) <- NULL


```

In the Poisson regression summarized in the second column of Table \@ref(tab:tab-gender-binary-count-ls), the outcome “positive words” is treated as a count variable rather than the binary presence/absence. Here we provide the rate ratios of the use of positive words (the exponential of the regression coefficient). These incident rate ratios essentially tell us how many more (or fewer) positive words are expected, compared to a reference for factor and binary variables, or for a unit change of the continuous explanatory variable. The interpretation of rate ratios for Poisson regression is closely related to that of odds ratios for logistic regression. No evidence for a gender effect on the count of positive words is found in any of the Poisson regression models for the different funding schemes. In Projects and Spark, younger researchers tend to use more positive words than their older peers. Unsurprisingly, more positive words can be found in longer texts. In Projects, researchers from the ETH Domain have a higher positive word count compared to cantonal universities, while in Careers, researchers affiliated in UAS and others tend to describe their research less positively than their peers from the cantonal universities.  


```{r tab-gender-count-ls, message = FALSE, warning = FALSE, cache = F, results = "asis"}
# regression results for reach research area
reg_gender_count_ls <- list(pf = pf_reg_ls$count, 
                            career = career_reg_ls$count,
                            spark = sparkabs_reg_ls$count)

ls_column_count <- lapply(reg_gender_count_ls,
                          FUN = tabulate_ci) %>% unlist()

var_column <- rep(c("Intercept", "gender - male", "age (years)",
                    "inst. type - ETH Domain", 
                    "inst. type - UAS and Other",
                    "text length (per 100 words)"), 3)

FI_column <- c(
  rep(paste0("Projects: ", format(nrow(pf_reg_ls$count$model), big.mark = "'"),
             " observations"), 6),
  rep(paste0("Careers: ", format(nrow(career_reg_ls$count$model),
                                 big.mark = "'"), " observations"), 6), 
  rep(paste0("Spark: ", format(nrow(sparkabs_reg_ls$count$model),
                               big.mark = "'"), " observations"), 6))

tab_count <- data.frame(FI_column, var_column, ls_column_count,
                        stringsAsFactors = FALSE)
row.names(tab_count) <- NULL

```


```{r tab-gender-binary-count-ls, message = FALSE, warning = FALSE, cache = F, results = "asis"}
tab_binary %>% 
  left_join(tab_count,
            by = c("FI_column", "var_column")) %>% 
  as_grouped_data(groups = "FI_column") %>%
  as_flextable(hide_grouplabel = TRUE,
               col_keys = c("var_column",
                            "ls_column", "ls_column_count")) %>% 
  colformat_md(j = c(2)) %>%
  set_header_labels(var_column = " ",
                    ls_column_count = "RR Estimate (95% CI)",
                    ls_column = "OR Estimate (95% CI)") %>%
  add_header_row(values = c(" ", "Logistic model for presence", "Poisson model for the count"), 
                 colwidths = c(1, 1, 1), top = TRUE) %>% 
  bold(bold = TRUE, part = c("header")) %>%
  set_caption(caption = "Summary of the logistic model using odds ratios (OR) and summary of the poisson model using rate ratios (RR), with 95% confidence intervals (CI). The reference for the gender variable is female and for the institution (inst.) type factor variable is 'Cantonal University'.") %>%
  width(width = 3) %>% 
  hline(i = c(7, 14)) %>% 
  bold(i = c(1, 8, 15))
```


```{r prep-fig-interaction-gender, cache = FALSE, results = "asis", warning = FALSE, message = FALSE}

get_predict_words <- function(m, freq_data){
  # For an applicant of average age with a proposal of average text 
  # length employed at a cantonal university
  newdata <- 
    tibble(ResponsibleApplicantGender = c("f", "m"),
           ResponsibleApplicantAgeAtSubmission = 
             rep(mean(freq_data$ResponsibleApplicantAgeAtSubmission), 2),
           ResearchInstitutionType = rep("Cantonal university", 2), 
           text_length100 = rep(mean(freq_data$text_length100), 2))
  
  p <- predict(m, newdata = newdata, type = "response", se.fit = TRUE)
  lwr <- with(p, fit + qnorm(0.025)*se.fit)
  upr <- with(p, fit + qnorm(0.975)*se.fit)
  
  return(tibble(ResponsibleApplicantGender = c("f", "m"),
                count = p$fit, 
                lwr = lwr, upr = upr))
}


spark_pred <- get_predict_words(sparkabs_reg_ls$count,
                                freq_data = sparkabs_freq_ls)

pf_pred <- get_predict_words(pf_reg_ls$count,
                             freq_data = pf_freq_ls)

career_pred <- get_predict_words(career_reg_ls$count,
                             freq_data = career_freq_ls)


fig_pf <- pf_pred %>% 
  mutate(ResponsibleApplicantGender = 
           case_when(ResponsibleApplicantGender == "f" ~ "Female",
                     ResponsibleApplicantGender == "m" ~ "Male")) %>% 
  ggplot(aes(ResponsibleApplicantGender, count)) +
  geom_pointrange(aes(ymin = lwr, ymax = upr),
                  position = position_dodge(width = 0.5)) +
  ylab("Estimated number of positive words") +
  xlab("Applicant gender") + 
  ggtitle("Projects") + 
  ylim(.8, 2) + 
  theme_minimal() + 
  theme(panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line(linetype = "dashed"),
        panel.grid.minor.y = element_line(linetype = "dashed")) 


fig_career <- career_pred %>% 
  mutate(ResponsibleApplicantGender = 
           case_when(ResponsibleApplicantGender == "f" ~ "Female",
                     ResponsibleApplicantGender == "m" ~ "Male")) %>% 
  ggplot(aes(ResponsibleApplicantGender, count))+
  geom_pointrange(aes(ymin = lwr, ymax = upr),
                  position = position_dodge(width = 0.5)) +
  ylab("Estimated number of positive words") +
  xlab("Applicant gender") + 
  ggtitle("Careers") +
  ylim(.8, 2) +
  theme_minimal() + 
  theme(axis.title.y = element_blank(),
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line(linetype = "dashed"),
        panel.grid.minor.y = element_line(linetype = "dashed")) 

fig_spark_abs <- spark_pred %>% 
  mutate(ResponsibleApplicantGender = 
           case_when(ResponsibleApplicantGender == "f" ~ "Female",
                     ResponsibleApplicantGender == "m" ~ "Male")) %>% 
  ggplot(aes(ResponsibleApplicantGender, count))+
  geom_pointrange(aes(ymin = lwr, ymax = upr),
                  position = position_dodge(width = 0.5)) +
  ylab("Estimated number of positive words") +
  xlab("Applicant gender") + 
  ggtitle("Spark") +
  ylim(.8, 2) +
  theme_minimal() + 
  theme(axis.title.y = element_blank(),
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line(linetype = "dashed"),
        panel.grid.minor.y = element_line(linetype = "dashed")) 
```

Figure \@ref(fig:fig-interaction-gender) illustrates the effects of gender on the language used in the Poisson regression models. The plots show the predicted number of positive words with 95% confidence error bars per gender, for an applicant of average age with a proposal of average text length from a cantonal university, for each funding scheme. 

We conclude that we cannot detect a difference in the use of positive words between male and female applicants in Life Sciences in any of the funding schemes. This is noteworthy, because the words used in our analysis were derived from PubMed abstracts (@Vinkers2015) and the original study (@Lerchenmueller) found a significant gender differences in their usage from clinical research and general life science articles indexed in PubMed. But that study only compared successful publications, while 
this analysis <!-- our study -->
includes both funded and rejected proposals.

<!-- Projects, the funding scheme with the largest sample size, shows gender differences in MINT and SSH. We observe evidence that male applicants in MINT and SSH use more positive words than their female counterparts. There is also weak evidence that male MINT applicants in Careers use more positive words than their female peers. -->

<!-- The trend is different for SSH applicants in Spark. In Spark, female applicants are more likely to use higher count of positive words in their texts than male applicants, unlike in Projects. -->


```{r fig-interaction-gender, fig.width = 6,fig.height = 3.5, fig.cap = "Illustration of the effects of gender on the language used in the Poisson regression models. The plots show the predicted number of positive words with 95% confidence error bars per gender, for an applicant of average age with a proposal of average text length, for each funding scheme. In all three funding schemes, male applicants used more positive words, but there was no significant difference as their confidence intervals overlap.", cache = FALSE, results = "asis", warning = FALSE, message = FALSE}

# Original caption: Illustration of the effects of gender and research area on the count of positive words for each funding instrument based on the fitted models. The plots are based on predicted number of positive used by applicants of each gender with average age and average abstract length of each instrument. In Spark, there is a small evidence that SSH female applicants used higher count of positive terms than their male peers, in constract to Projects, where male applicants are likelier to use higher count of positive words.

ggarrange(fig_pf, fig_career, fig_spark_abs,
          nrow = 1)
```


# Discussion

Our interest in this topic was motivated in part by the desire to understand if the grant evaluation process is influenced by language use.  If it is, this could be contributing to the “leaky pipeline” of women leaving academic careers. Our results indicate that the use of self-promotional language – at least to the limited extent analyzed in this study – does not appear to differ significantly between male and female grant applicants.

The SNSF first looked into the use of self-promotional language after the first Spark call, when it was found that, in spite of the double-blind evaluation process, female applicants had a significantly lower success rate than their male counterparts. One hypothesis was that perhaps the proposals themselves were written in different styles. However, our analysis here shows that there is no evidence for a significant difference in the use of positive words between the proposal texts of men and women. We performed a separate statistical analysis, beyond the scope of the current work, that looked at several confounding variables, and we found that adjusting for the confounding variables (discipline, type of research institution, applicant age) explained the observed differences in success rates between the genders. In the second Spark funding call, no evidence for a difference in success rates between men and women could be observed.  _Could remove this paragraph_

In Projects, our gender analysis is limited by the fact that we only consider the gender of the corresponding applicant, even though proposals are often submitted by multiple applicants. While it has historically been the case that corresponding applicants generally write most of the proposal, we recognize that many Projects proposals are written collaboratively. The SNSF stands firm in its belief that collaborative research is very important for scientific research.

The 25 "positive words" used in our analysis were derived from life sciences abstracts found in PubMed (@Vinkers2015). Similar to the original @Lerchenmueller study, we found that successful life sciences Projects proposals were more likely to contain these positive words, although we found no evidence for such an effect in the Careers or Spark funding schemes.  It is possible that the observed effect of positive words in Projects but not in the other two funding schemes may simply be due to its higher sample size, and consequently higher statistical power.  However, it is interesting to note that no affect for any gender difference was found when considering both successful and rejected proposals together, even though this is a much bigger data set.  The direction of the effect also changed in Careers and Spark between the full data set and only the successful data set, although the effect was not significantly significant.

This work is a replication study that analyses the difference in the use of positive language between male and female grant applicants.  The analysis presented here must not be interpreted as providing any evidence of the effect of positive language on grant funding decisions.  Many other confounding variables affect grant funding decisions, and we intentionally analyze all proposals in an entirely separate analysis from successful proposals.  A more complete analysis of all relevant variables that go into grant funding decisions is left for future work.

Overall, we see it as an encouraging result that we do not observe any large differences in the use of self-promotional language between the genders in their grant proposals.


# Acknowledgements

The authors would like to thank Fabio Molo for his significant contributions to the inception and development of this project.

# Author contributions (CRediT)
Conceptualization: RH, AJ; Data curation: SL, RH; Formal Analysis: SL, RH;  Investigation: RH, SL, AJ; Methodology: SL, RH; Project administration: SL, RH; Resources: RH, AJ; Software: SL, RH; Supervision: RH, AJ; Validation:  RH, AJ; Visualization: SL, RH; Writing – original draft: RH, SL; Writing – review & editing: RH, SL, AJ.

# Competing interests
SL and RH were employees of the Swiss National Science Foundation during the initial analysis leading to this manuscript. AJ is still employed at the SNSF.

# Data availability statement

The raw data in this study includes the titles and abstracts of rejected grant proposals, which are impossible to anonymize and considered confidential by the SNSF. The aggerated data including the counts of positive words in the texts was deposited on Zenodo and is available through 
[10.5281/zenodo.6676561](https://zenodo.org/records/6676562). The `R` scripts used to compile this manuscript are available from github,  [github.com/snsf-data/positive_language](https://github.com/snsf-data/positive_language).

# References

<div id="refs"></div>

